---
title: "Final Lab Gas Data Processing"
author: "Nibia Becerra Santillan"
format:
  html:
    toc: true
    code-fold: true
execute:
  echo: true
  warning: false
  message: false
categories: [R]
description: "Pipeline for lab gas fluxes with tidyverse and QC checks."
date: 2025-04-29
---

```{r}
library(forcats)
library(tidyverse)
library(lme4)
library(emmeans)
library(scales)
```

# Reading data

```{r}
fluxData <- read_csv("./combo_FINAL_FLUXES_gSOIL.csv")
```

# Treatment assignmnet from JarID

### Evaluating which grouping was most appropiate due to sampling size differences throught the experimental stages

```{r}
fluxData_treatment_three_jars <- fluxData %>%
  mutate(
    treatment = case_when(
      JarID == "BLANK" ~ "BLANK",
      between(as.numeric(JarID), 1, 3) ~ "90-25",
      between(as.numeric(JarID), 7, 9) ~ "50-25",
      between(as.numeric(JarID), 13, 15) ~ "20-25",
      between(as.numeric(JarID), 19, 21) ~ "90-35",
      between(as.numeric(JarID), 25, 27) ~ "50-35",
      between(as.numeric(JarID), 31, 33) ~ "20-35",
      TRUE ~ NA_character_
    )) 

fluxData_treatment_six_jars <- fluxData %>%
  mutate(
    treatment = case_when(
      JarID == "BLANK" ~ "BLANK",
      between(as.numeric(JarID), 1, 6) ~ "90-25",
      between(as.numeric(JarID), 7, 12) ~ "50-25",
      between(as.numeric(JarID), 13, 18) ~ "20-25",
      between(as.numeric(JarID), 19, 24) ~ "90-35",
      between(as.numeric(JarID), 25, 30) ~ "50-35",
      between(as.numeric(JarID), 31, 36) ~ "20-35",
      TRUE ~ NA_character_ 
    )
  )
```

# Fix Dates

```{r}

# Three jars
flux_three_jars <- fluxData_treatment_three_jars  %>% 
  mutate(Dateflux = fluxData_treatment_three_jars$Date) %>% 
  mutate(Date_corr = mdy(Dateflux))

# Six jars
flux_six_jars <- fluxData_treatment_six_jars %>% 
  mutate(Dateflux = fluxData_treatment_six_jars$Date) %>% 
  mutate(Date_corr = mdy(Dateflux))

```

# Means of Fluxes No WET UP

```{r}

NO_WETUP_six <- flux_six_jars%>% 
  filter(Date_corr %in% c("2024-07-16", "2024-07-17", "2024-07-20", "2024-07-22", "2024-07-23")) %>% 
  group_by(treatment, Date_corr) %>% 
  summarize(
    mean_CO2 = mean(finalC02flux_gSoil_hr),
    sd_CO2 = sd(finalC02flux_gSoil_hr),
    se_CO2 = sd_CO2 / sqrt(n()),  
    mean_CH4 = mean(finalCH4flux_gSoil_hr),
    sd_CH4 = sd(finalCH4flux_gSoil_hr),
    se_CH4 = sd_CH4 / sqrt(n()),  
    .groups = "drop") %>% 
    separate(treatment, into = c("WHC", "Temperature"), remove = FALSE) %>%
  mutate(
    WHC = factor(WHC, levels = c("20", "50", "90")),
    Temperature = factor(Temperature, levels = c("25", "35")))
```

# Colors and lines

```{r}
graph_colors <- c("20" = "#E69F00", "50" = "#5290bf", "90" = "#08306B", "NA"= "grey3", "25" = "#5290bf", "35" = "#E69F00")
types <- c("25" = "solid", "35" = "longdash", "30" = "dotted", "NA"= "solid")
```

# Antecedent Data

# CO2

# Timeseries Graph

```{r}
NO_WETUP_C02_timeseries <- ggplot(NO_WETUP_six, 
       aes(x = Date_corr, y = mean_CO2, 
           group = treatment,
           color = WHC,
           linetype = Temperature)) +
  geom_rect(aes(xmin = as.Date("2024-07-16") - 0.5, xmax = as.Date("2024-07-17") + 0.5 , ymin =  0, ymax = Inf), 
            fill = "#e6ecef", color = NA) +
  geom_rect(aes(xmin = as.Date("2024-07-20") - 0.5, xmax = as.Date("2024-07-23") + 0.5, ymin = 0, ymax = Inf), 
            fill = "#efe8e6", color = NA) +
  geom_point() +
  geom_line()+
  geom_errorbar(aes(ymin = mean_CO2 - se_CO2, 
                    ymax = mean_CO2 + se_CO2,
                    color = WHC,
                    linetype = Temperature),
                width = 0.5) +
  scale_color_manual(values = graph_colors, name = "WHC (%)") +
  scale_linetype_manual(values = types, name = "Temperature") +
  scale_x_date(breaks = unique(NO_WETUP_six$Date_corr), 
               date_labels = "%d %b\n%Y") +
  labs(
    title = "CO₂ Flux Over Time by Treatment",
    x = "Date", 
    y = "Mean CO₂ Flux ± SE"
  ) +
  theme_classic() +
  theme(
    legend.position = "right",
    plot.title = element_text(hjust = 0.5, face = "bold"),
    axis.title = element_text(face = "bold"),
    strip.text = element_text(face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1)  
  ) +
  annotate("text", x = as.Date("2024-07-16") + 0.5, y = 2.3e-8, label = "Only moisture", color = "#4d4c4c", fontface = "bold", size = 3) +
  annotate("text", x = as.Date("2024-07-20") + 1.5, y = 2.3e-8, label = "Moisture and Temperature", color = "#4d4c4c", fontface = "bold", , size = 3)


NO_WETUP_C02_timeseries

# Commented out for sharing purposes
# ggsave(
#   filename = "LAB_C02_ActualData.png",
#   plot = NO_WETUP_C02_timeseries,
#   path = "~/Desktop/2024-2025/Spring 2025/Honors_Spring/Code/cso066code_BecerraSantillanHonors/TRACEincubation-Figures/Gas",
#   width = 8,    # adjust width as needed
#   height = 6,   # adjust height as needed
#   dpi = 300     # adjust dpi for resolution
# )
```

```{r}

NO_WETUP_C02_timeseries <- NO_WETUP_six %>% 
  filter(WHC != "NA", Temperature != "NA") %>% 
mutate(
  Date_corr = as.Date(Date_corr),
  Hours = case_when(
    Date_corr == as.Date("2024-07-16") ~ 25,
    Date_corr == as.Date("2024-07-17") ~ 48,
    Date_corr == as.Date("2024-07-20") ~ 120,
    Date_corr == as.Date("2024-07-22") ~ 168,
    Date_corr == as.Date("2024-07-23") ~ 192
  )) %>% 
  ggplot( 
       aes(x = Hours, y = mean_CO2, 
           group = treatment,
           color = Temperature)) +
  geom_point(alpha = 10) +
  geom_line(linewidth = 1.5)+
  geom_errorbar(aes(ymin = mean_CO2 - se_CO2, 
                    ymax = mean_CO2 + se_CO2,
                    color = Temperature),
                width = 5) +
  scale_color_manual(values = graph_colors, name = "Temperature") +
  scale_x_continuous(
  breaks = c(25, 48, 120, 168, 192),
  labels = c("25 \n\ hrs", "48 \n\ hrs", "120 \n\ hrs", "168 \n\ hrs", "192 \n\ hrs")) +
  labs(
    x = "Incubation Hours", 
    y = "Mean CO₂ Flux ± SE (µmol g -¹ h -¹ )"
  ) +
  theme_classic() +
  theme(
    legend.direction = "horizontal", 
    legend.box.just = "center", 
    legend.position = c(0.18, 0.80),  
    legend.background = element_blank(),
    legend.box.background = element_rect(color = "transparent"),
    legend.key = element_blank(),
    plot.title = element_text(hjust = 0.5, face = "bold"),
    axis.title = element_text(face = "bold"),
    strip.text = element_text(face = "bold"),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    strip.background = element_rect(color = "black", fill = "white", linewidth = 1)) +
  facet_grid(.~WHC)


NO_WETUP_C02_timeseries

# Commented out for sharing purposes
ggsave(
  filename = "LAB_C02_Facet_grid_WHC.png",
  plot = NO_WETUP_C02_timeseries,
  path = "~/Desktop/2024-2025/",
  width = 12,    
  height = 6,   
  dpi = 300     
)
```

# Mixed Model effect

## Making "Phase" groups for jars

```{r}
flux_six_jars  <- flux_six_jars  %>%
  mutate(phase = ifelse(Date_corr %in% as.Date(c("2024-07-16", "2024-07-17")), "MoistureOnly", "Moisture+Temp")) %>% 
  mutate(phase = as.factor(phase))
```

```{r}
m4_phase_CO2 <- lmer(finalC02_flux ~ treatment * phase + (1|JarID), 
                 data = flux_six_jars, 
                 na.action = na.omit)
summary(m4_phase_CO2)


# Interactions

emm_CO2 <- emmeans(m4_phase_CO2, ~ treatment * phase)
contrast(emm_CO2, interaction = TRUE, adjust = "sidak")

```

#CH4

# Timeseries graph

```{r}
# All in the same plot
NO_WETUP_CH4_timeseries <- ggplot(NO_WETUP_six, 
       aes(x = Date_corr, y = mean_CH4, 
           group = treatment,
           color = WHC,
           linetype = Temperature)) +
  geom_rect(aes(xmin = as.Date("2024-07-16") - 0.5, xmax = as.Date("2024-07-17") + 0.5 , ymin =  -1e-9, ymax = Inf), 
            fill = "#e6ecef", color = NA) +
  geom_rect(aes(xmin = as.Date("2024-07-20") - 0.5, xmax = as.Date("2024-07-23") + 0.5, ymin = -1e-9, ymax = Inf), 
            fill = "#efe8e6", color = NA) +
  geom_point()+
  geom_line() +
  geom_errorbar(aes(ymin = mean_CH4 - sd_CH4, 
                    ymax = mean_CH4 + sd_CH4,
                    color = WHC,
                    linetype = Temperature),
                width = 0.5) +
  scale_color_manual(values = graph_colors, name = "WHC (%)") +
  scale_linetype_manual(values = types, name = "Temperature") +
  scale_x_date(breaks = unique(NO_WETUP_six$Date_corr), 
               date_labels = "%d %b\n%Y") + 
  labs(
    title = "CH4 Flux Over Time by Treatment",
    x = "Date", 
    y = "Mean CH4 Flux ± SD"
  ) +
  theme_classic() +
  theme(
    legend.position = "right",
    plot.title = element_text(hjust = 0.5, face = "bold"),
    axis.title = element_text(face = "bold"),
    strip.text = element_text(face = "bold")
  )+
  annotate("text", x = as.Date("2024-07-16") + 0.5, y = 7.5e-9 , label = "Only Moisture", color = "#4d4c4c", fontface = "bold", size = 3.5) +
  annotate("text", x = as.Date("2024-07-20")+ 1.5 , y = 7.5e-9, label = "Moisture and Temperature", color = "#4d4c4c", fontface = "bold", size = 3.5)


NO_WETUP_CH4_timeseries

# ggsave(
#   filename = "LAB_CH4_ActualData.png",
#   plot = NO_WETUP_CH4_timeseries,
#   path = "~/Desktop/2024-2025/Spring 2025/Honors_Spring/Code/cso066code_BecerraSantillanHonors/TRACEincubation-Figures/Gas",
#   width = 8,    # adjust width as needed
#   height = 6,   # adjust height as needed
#   dpi = 300     # adjust dpi for resolution
# )

```

```{r}
# Different graphs 
NO_WETUP_CH4_timeseries_facet <- ggplot(NO_WETUP_six, 
       aes(x = Date_corr, y = mean_CH4, 
           group = treatment,
           color = WHC,
           linetype = Temperature)) +
  geom_rect(aes(xmin = as.Date("2024-07-16") - 0.5, xmax = as.Date("2024-07-17") + 0.5 , ymin =  -1e-10, ymax = Inf), 
            fill = "#e6ecef", color = NA) +
  geom_rect(aes(xmin = as.Date("2024-07-20") - 0.5, xmax = as.Date("2024-07-23") + 0.5, ymin = -1e-10, ymax = Inf), 
            fill = "#efe8e6", color = NA) +
  geom_point() +
  geom_line()+
  geom_errorbar(aes(ymin = mean_CH4 - sd_CH4, 
                    ymax = mean_CH4 + sd_CH4,
                    color = WHC,
                    linetype = Temperature),
                width = 0.5) +
  scale_color_manual(values = graph_colors, name = "WHC (%)") +
  scale_linetype_manual(values = types, name = "Temperature") +
  labs(
    title = "CH4 Flux Over Time by Treatment",
    x = "Date", 
    y = "Mean CH4 Flux ± SD"
  ) +
  theme_classic() +
  theme(
    legend.position = "right",
    plot.title = element_text(hjust = 0.5, face = "bold"),
    axis.title = element_text(face = "bold"),
    strip.text = element_text(face = "bold")
  ) +
  facet_wrap(~ treatment, scales = "free_y")
NO_WETUP_CH4_timeseries_facet


# ggsave(
#   filename = "LAB_CH4_ActualData_facet.png",
#   plot = NO_WETUP_CH4_timeseries_facet,
#   path = "~/Desktop/2024-2025/Spring 2025/Honors_Spring/Code/cso066code_BecerraSantillanHonors/TRACEincubation-Figures/Gas",
#   width = 8,    # adjust width as needed
#   height = 6,   # adjust height as needed
#   dpi = 300     # adjust dpi for resolution
# )
```

## Models for Phases

```{r}
m4_phase_CH4 <- lmer(finalCH4_flux ~ treatment * phase + (1|JarID), 
                 data = flux_six_jars, 
                 na.action = na.omit)
summary(m4_phase_CH4)

emm_CH4 <- emmeans(m4_phase_CH4, ~ treatment * phase)
contrast(emm_CH4, interaction = TRUE, adjust = "sidak")

```

# WETUP

## Adding a wet up stage column

```{r}
wet_up_cols <- c("PRE_WET", "two_min", "one_hour", "four_hour", "eight_hour", "twenty_four_hour")

#flux_three_jars

WETUP <- flux_three_jars %>%
  mutate(across(all_of(wet_up_cols), ~ na_if(., "NA"))) %>%
  rowwise() %>%
  mutate(
    WET_UP_stage = {
      values <- c_across(all_of(wet_up_cols))
      cols <- wet_up_cols[!is.na(values)]
      ifelse(length(cols) > 0, str_c(cols, collapse = ", "), NA)
    }
  ) %>%
  ungroup() %>%
  mutate(
    WET_UP_stage = factor(
      WET_UP_stage,
      levels = wet_up_cols,
      ordered = TRUE
    )
  )

WETUP
```

## Filtering

```{r}

WETUP <- WETUP %>% 
  select(Model, JarID, Time, Label, finalC02_flux, finalCH4_flux, treatment,Dateflux,Date_corr, WET_UP_stage) %>% 
  filter(Date_corr %in% c("2024-07-29", "2024-07-30")) %>% 
  separate(treatment, into = c("WHC", "Temperature"), remove = FALSE) %>%
  mutate(
    WHC = factor(WHC, levels = c("20", "50", "90")),
    Temperature = factor(Temperature, levels = c("25", "35", "30")),
    WET_UP_stage_correct = case_when(
      WET_UP_stage == "PRE_WET" ~ 8,
      WET_UP_stage == "eight_hour" ~ 0,
      WET_UP_stage == "two_min" ~ 2,
      WET_UP_stage == "one_hour" ~ 1,
      WET_UP_stage == "four_hour" ~ 4,
      WET_UP_stage == "tweny_four_hour" ~ 24
    )) 
#WETUP_summ


WETUP_summ <- WETUP %>% 
  group_by(treatment, WET_UP_stage_correct) %>%  
  summarize(
    mean_CO2 = mean(finalC02_flux),
    sd_CO2 = sd(finalC02_flux),
    se_CO2 = sd_CO2 / sqrt(n()),  
    mean_CH4 = mean(finalCH4_flux),
    sd_CH4 = sd(finalCH4_flux),
    se_CH4 = sd_CH4 / sqrt(n()), 
    .groups = "drop"
  ) %>% 
  separate(treatment, into = c("WHC", "Temperature"), remove = FALSE) %>%
  mutate(
    WHC = factor(WHC, levels = c("20", "50", "90")),
    Temperature = factor(Temperature, levels = c("25", "35", "30")))
#WETUP_summ

```

## CO2 WET UP Graph

```{r}
# CO2

WETUP_ActualMeasurements_CO2 <- WETUP_summ %>% 
  ggplot(aes(x = WET_UP_stage_correct, y = mean_CO2,
             group = treatment,
             color = WHC,
             linetype = Temperature)) +
  geom_point() +
  geom_line()+
  geom_errorbar(
    aes(ymin = mean_CO2 - se_CO2, ymax = mean_CO2 + se_CO2),
    width = 0.5
  ) +
  scale_color_manual(values = graph_colors, name = "WHC (%)") +
  scale_linetype_manual(values = types, name = "Temperature") +
  scale_x_continuous(
  breaks = c(0,2,1,4,8,24),
  labels = c("0 min", "2 min", "1 hr", "4 hrs", "8 hrs", "24 hrs"))+
  labs(
    x = "Time since Wet Up", 
    y = "Mean CO₂ Flux ± SE"
  ) +
  theme_classic() +
  theme(
    legend.position = "right",
    plot.title = element_text(hjust = 0.5, face = "bold"),
    axis.title = element_text(face = "bold"),
    strip.text = element_text(face = "bold")
  ) 

WETUP_ActualMeasurements_CO2 

ggsave(
  filename = "LAB_WETUP_ActualMeasurements_CO2.png",
  plot = WETUP_ActualMeasurements_CO2,
  path = "~/Desktop/2024-2025/",
  width = 12,   
  height = 6,   
  dpi = 300     
)
```
# Temperature facets
```{r}
# CO2

WETUP_ActualMeasurements_CO2_facets_temp_log <- WETUP_summ %>% 
  filter(Temperature != "NA") %>% 
  ggplot(aes(x = WET_UP_stage_correct, y = mean_CO2,
             group = treatment,
             color = Temperature)) +
  geom_point() +
  geom_line() +
  geom_errorbar(
    aes(ymin = mean_CO2 - se_CO2, ymax = mean_CO2 + se_CO2),
    width = 1
  ) +
  scale_color_manual(values = graph_colors, name = "Temperature") +
  scale_x_continuous(
  breaks = c(0,2,1,4,8,24),
  labels = c("0 min", "2 min", "1 hr", "4 hrs", "8 hrs", "24 hrs"))+
  scale_y_log10() +
  labs(
    x = "Time since Wet Up", 
    y = "Mean CO₂ Flux ± SE"
  ) +
  theme_classic() +
  theme(
    legend.position = "right",
    plot.title = element_text(hjust = 0.5, face = "bold"),
    axis.title = element_text(face = "bold"),
    strip.text = element_text(face = "bold"),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1)
  ) +
  facet_wrap(~WHC)
WETUP_ActualMeasurements_CO2_facets_temp_log

ggsave(
  filename = "WETUP_ActualMeasurements_CO2_facets_temp_log.png",
  plot = WETUP_ActualMeasurements_CO2_facets_temp_log,
  path = "~/Desktop/2024-2025/",
  width = 12,    
  height = 6,   
  dpi = 300    
)
  
```
## WHC facets 
```{r}
# CO2

WETUP_ActualMeasurements_CO2_facets_WHC_log <- WETUP_summ %>% 
  filter(Temperature != "NA") %>% 
  ggplot(aes(x = WET_UP_stage_correct, y = mean_CO2,
             group = treatment,
             color = WHC)) +
  geom_point() +
  geom_line() +
  geom_errorbar(
    aes(ymin = mean_CO2 - se_CO2, ymax = mean_CO2 + se_CO2),
    width = 1
  ) +
  scale_color_manual(values = graph_colors, name = "Water Holding Capacity") +
  scale_x_continuous(
  breaks = c(0,2,1,4,8,24),
  labels = c("0 min", "2 min", "1 hr", "4 hrs", "8 hrs", "24 hrs"))+
  scale_y_log10() +
  labs(
    x = "Time since Wet Up", 
    y = "Mean CO₂ Flux ± SE"
  ) +
  theme_classic() +
  theme(
    legend.position = "right",
    plot.title = element_text(hjust = 0.5, face = "bold"),
    axis.title = element_text(face = "bold"),
    strip.text = element_text(face = "bold"),
     panel.border = element_rect(color = "black", fill = NA, linewidth = 1)
  ) +
  facet_wrap(~Temperature)
WETUP_ActualMeasurements_CO2_facets_WHC_log

ggsave(
  filename = "WETUP_ActualMeasurements_CO2_facets_WHC_log.png",
  plot = WETUP_ActualMeasurements_CO2_facets_WHC_log,
  path = "~/Desktop/2024-2025/",
  width = 12,    
  height = 6,   
  dpi = 300    
)
  
```

# Mixed effects model

```{r}

WETUP_no_PRE <- WETUP %>% 
  filter(WET_UP_stage != "PRE_WET") #PRE_WET are the conditions before the experiement was applied, so we wanted to remove it from the model of "recovery"

WETUP_CO2 <- lmer(finalC02_flux ~ WHC * Temperature * WET_UP_stage + (1|JarID), 
                 data =  WETUP_no_PRE, 
                 na.action = na.omit)
summary(WETUP_CO2)

WETUP_emm_CO2 <- emmeans(WETUP_CO2, ~ WHC * Temperature * WET_UP_stage)
contrast(WETUP_emm_CO2, interaction = TRUE, adjust = "sidak")

```

## CH4

```{r}
# WETUP_ActualMeasurements_CH4 <- WETUP_summ %>% 
#   ggplot(aes(x = WET_UP_stage, y = mean_CH4,
#              group = treatment,
#              color = WHC,
#              linetype = Temperature)) +
#   geom_point() +
#   geom_line()+
#   geom_errorbar(
#     aes(ymin = mean_CH4 - se_CH4, ymax = mean_CH4 + se_CH4),
#     width = 0.5
#   ) +
#   scale_color_manual(values = graph_colors, name = "WHC (%)") +
#   scale_linetype_manual(values = types, name = "Temperature") +
#   scale_x_discrete( labels = c("0 min", "3 min", "1 hr", "4 hr", "8 hr", "24 hr") )+
#   labs(
#     title = "WET-UP CH₄ Flux Over Time by Treatment",
#     x = "Stage", 
#     y = "Mean CH₄ Flux ± SE"
#   ) +
#   theme_classic() +
#   theme(
#     legend.position = "right",
#     plot.title = element_text(hjust = 0.5, face = "bold"),
#     axis.title = element_text(face = "bold"),
#     strip.text = element_text(face = "bold")
#   ) 
# 
# WETUP_ActualMeasurements_CH4

# ggsave(
#   filename = "LAB_WETUP_ActualMeasurements_CH4.png",
#   plot = WETUP_ActualMeasurements_CH4,
#   path = "~/Desktop/2024-2025/Spring 2025/Honors_Spring/Code/cso066code_BecerraSantillanHonors/TRACEincubation-Figures/Gas",
#   width = 8,    # adjust width as needed
#   height = 6,   # adjust height as needed
#   dpi = 300     # adjust dpi for resolution
# )
```

```{r}

# WETUP_ActualMeasurements_CH4_facet <- WETUP_summ %>% 
#   ggplot(aes(x = WET_UP_stage, y = mean_CH4,
#              group = treatment,
#              color = WHC,
#              linetype = Temperature)) +
#   geom_point() +
#   geom_line()+
#   geom_errorbar(
#     aes(ymin = mean_CH4 - se_CH4, ymax = mean_CH4 + se_CH4),
#     width = 0.5
#   ) +
#   scale_color_manual(values = graph_colors, name = "WHC (%)") +
#   scale_linetype_manual(values = types, name = "Temperature") +
#   scale_x_discrete( labels = c("0 min", "3 min", "1 hr", "4 hr", "8 hr", "24 hr") )+
#   labs(
#     title = "WET-UP CH₄ Flux Over Time by Treatment",
#     x = "Stage", 
#     y = "Mean CH₄ Flux ± SE"
#   ) +
#   theme_classic() +
#   theme(
#     legend.position = "right",
#     plot.title = element_text(hjust = 0.5, face = "bold"),
#     axis.title = element_text(face = "bold"),
#     strip.text = element_text(face = "bold")
#   ) +
#   facet_wrap(~treatment, scales = "free_y")
# 
# WETUP_ActualMeasurements_CH4_facet
# 
# # ggsave(
# #   filename = "LAB_WETUP_ActualMeasurements_CH4_facet.png",
# #   plot = WETUP_ActualMeasurements_CH4_facet,
# #   path = "~/Desktop/2024-2025/Spring 2025/Honors_Spring/Code/cso066code_BecerraSantillanHonors/TRACEincubation-Figures/Gas",
# #   width = 8,    # adjust width as needed
# #   height = 6,   # adjust height as needed
# #   dpi = 300     # adjust dpi for resolution
# # )

```

```{r}
WETUP_CH4 <- lmer(finalCH4_flux ~ WHC * Temperature * WET_UP_stage + (1|JarID), 
                 data =  WETUP_no_PRE, 
                 na.action = na.omit)
summary(WETUP_CH4)

WETUP_emm_CH4 <- emmeans(WETUP_CH4, ~ WHC * Temperature * WET_UP_stage)
contrast(WETUP_emm_CH4, interaction = TRUE, adjust = "sidak")

```
